---
title: "chapter_8"
author: "Jrose"
date: '2022-09-16'
output: html_document
---

# Chapter 8: Feature Engineering with recipes

Feature engineering involves reformatting predictors to make them easier for a model to use effectively

You also may need to reformat in order to use a certain model. An appendix table of preprocessing required for different model types can be found here [<https://www.tmwr.org/pre-proc-table.html#pre-proc-table>]

### 8.1 A simple recipe() for the AMES housing data

AMES data setup from section 7.7

```{r}
library(tidymodels)
tidymodels_prefer()

data(ames)

ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split) # Includes the recipes package

lm_model <- 
  linear_reg() %>% 
  set_engine("lm")

lm_wflow <- 
  workflow() %>% 
  add_model(lm_model) %>%
  add_formula(Sale_Price ~ Longitude + Latitude)   %>%
  remove_formula() %>% 
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))

```

Use of recipe

```{r}
simple_ames <- 
  recipe(Sale_Price ~ Neighborhood + Gr_Liv_Area + Year_Built + Bldg_Type,
         data = ames_train) %>%
  step_log(Gr_Liv_Area, base = 10) %>% 
  step_dummy(all_nominal_predictors())
simple_ames
```

In `recipe()` you can define a formula and the dataset

The rest of the feature engineering is done using "step_" functions

step_log() converts Gr_Liv_Area to log

step_dummy() converts qualitative data to a quantitative format (i.e. encodes 1 or 0s)

all_nominal_predictors() recognizes and captures the names of any predictor column that is currently a factor or character (tidy_select-like function)

Also available:

* all_numeric_predictors()
* all_numeric()
* all_predictors()
* all_outcomes()

Advantages of using recipes/steps:

* Recycling across models
* Recipes have more options for preprocessing vs add_formula
* Tidy_select variable selection
* All data processing in a single R object

### 8.2 Using Recipes

Let's add this to a workflow like in chapter 7

```{r}
lm_wflow <- 
  lm_wflow %>% 
  remove_variables() %>% 
  #^Need to remove variables added earlier (chap 7) before adding new recipe
  add_recipe(simple_ames)
lm_wflow
```

```{r}
lm_fit <- fit(lm_wflow, ames_train)
predict(lm_fit, ames_test %>% slice(1:3))
```


```{r}
lm_fit %>% 
  extract_recipe(estimated = TRUE)

# To tidy the model fit: 
lm_fit %>% 
  # This returns the parsnip object:
  extract_fit_parsnip() %>% 
  # Now tidy the linear model object:
  tidy() %>% 
  slice(1:5)
```

## 8.3 How Data are used by the recipe()


