---
title: "chapter_7"
author: "Jrose"
date: '2022-09-15'
output: html_document
---

# Chapter 7: A Model Workflow

Workflows for different types of models can be different.

They involve more than just fitting i.e.:

-   Excluding predictors
-   Feature selection
-   Imputing missing predictors
-   Feature transformation

Workflow = pipeline (general term, not %\>%)

```{r}
library(tidymodels)  # Includes the workflows package
tidymodels_prefer()

data(ames)
ames <- ames %>% mutate(Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)

lm_model <- 
  linear_reg() %>% 
  set_engine("lm")
```

### 7.2 Workflow Basics

Start with adding a model to a workflow

```{r}
lm_wflow <- 
  workflow() %>% 
  add_model(lm_model)

lm_wflow
```

Preprocessing has not been set yet. We can do a very simple one using an R formula

```{r}
lm_wflow <- 
  lm_wflow %>% 
  add_formula(Sale_Price ~ Longitude + Latitude)

lm_wflow
```

Fit using the fit() function

```{r}
lm_fit <- fit(lm_wflow, ames_train)
lm_fit
```

And predict

```{r}
predict(lm_fit, ames_test %>% slice(1:3))
```

Model and preprocessor can both be modified

```{r}
lm_fit %>% update_formula(Sale_Price ~ Longitude)
```

### 7.3 Adding raw variables to the workflow()

The `add_variables()` function uses tidyselect-like approach to select variables

```{r}
lm_wflow <- 
  lm_wflow %>% 
  remove_formula() %>% 
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))
lm_wflow
```

Other options for selecting predictors:

`predictors = c(ends_with("tude"))`

`predictors = everything()`

When fit, the workflow will assemble these data into a dataframe and pass it to the underlying function

```{r}
fit(lm_wflow, ames_train)
```

### 7.4 How does a workflow() use the formula?

There are a lot of different ways that underlying packages use the formula notation in R. Workflow() tries to unify them a little

Some models accept special notation for random effects

`lmer(distance~ Sex + (age|Subject), data=Orthodont)`

The (age\|Subject) notation says that each subject will have an estimated intercept and slope parameter for age

Used in a workflow:

```{r}
library(multilevelmod)


multilevel_spec <- linear_reg() %>% set_engine("lmer")

multilevel_workflow <- 
  workflow() %>% 
  # Pass the data along as-is: 
  add_variables(outcome = distance, predictors = c(Sex, age, Subject)) %>% 
  add_model(multilevel_spec, 
            # This formula is given to the model
            formula = distance ~ Sex + (age | Subject))

multilevel_fit <- fit(multilevel_workflow, data = Orthodont)
multilevel_fit
```

Basically workflow accepts package-specific formula syntax depending on what model is included

### 7.5 Creating Multiple Workflows at Once

Sometimes you have to try multiple different models on a set of data to find which works best.

-   Predictive models require evaluating a variety of different model types

-   Sequential testing of models typically starts with a "full model" using all predictors which is compared to a sequence of models removed each predictor in turn. Allows you to isolate the effects of each predictor

Because of this you can create a set of workflows

```{r}
location <- list(
  longitude = Sale_Price ~ Longitude,
  latitude = Sale_Price ~ Latitude,
  coords = Sale_Price ~ Longitude + Latitude,
  neighborhood = Sale_Price ~ Neighborhood
)

library(workflowsets)
location_models <- workflow_set(preproc = location, models = list(lm = lm_model))
location_models
```

```{r}
location_models$info[[1]]
extract_workflow(location_models, id = "coords_lm")
```

Workflow sets are mostly designed to work with resampling (discussed in Chapter 10)

You can fit each model in a workflow set using map

```{r}
location_models <-
   location_models %>%
   mutate(fit = map(info, ~ fit(.x$workflow[[1]], ames_train)))
location_models
```

An easier way will be introduced in Chap 11.1

### 7.6 Evaluating the Test Set

`last_fit()` fits the model fed to it to the entire training set and evaluates it using the testing set

```{r}
final_lm_res <- last_fit(lm_wflow, ames_split)
#^Done on single workflow, not set!
#^Also not it takes the split object
final_lm_res
```

Extract the fitted workflow

```{r}
fitted_lm_wflow <- extract_workflow(final_lm_res)
```

`collect_metrics()` and `collect_predictions()` provide access to the performance and predictions

```{r}
collect_metrics(final_lm_res)
collect_predictions(final_lm_res) %>% slice(1:5)
```

### 7.7 Chapter Summary

-   Modeling process involves more than just fitting (preprocessing, transformations, etc)

-   Model workflows capture the important components of the modeling process

-   Multiple workflows can be created inside of workflow sets

-   The `last_fit()` function fits final model to training and evaluates on test set
