---
title: "Chapter_6"
author: "Jrose"
date: '2022-04-03'
output: html_document
---

# Chapter 6: Fitting Models with parsnip

Getting used to the parsnip package from within tidymodels

### 6.1 Create a model

Some thoughts on linear regression models first:

-   Ordinary linear regression used least squares to solve for paremeters

    -   You would use `lm(formula, data, ...)`

-   Regularized linear regression adds penalty to least squares to encourage simplicity by removing predictors or shrinking coeffcients

    -   For bayesian: Use `stan_glm(formula, data, family = "gaussian", …)`

    -   For non-Bayesian: Use `glmnet(x=matrix, y=vector, family = "gaussian", …)`

Each has a different syntax for parameters and inputs

The niche of tidymodels is to unify approaches by asking you to:

1.  Specify the type of model (i.e. linear regression, random forest, etc)
2.  Specify the engine for fitting (i.e. what package is used...glmnet, etc)
3.  When required, specify the mode (is the outcome numeric, categorical, etc)

```{r}
library(tidymodels)
tidymodels_prefer()

linear_reg() %>% set_engine("lm")

linear_reg() %>% set_engine("glmnet") 

linear_reg() %>% set_engine("stan")
```

> "The parsnip package allows the user to be indifferent to the interface of the underlying model"

`translate()` shows how the data is fit into the underlying functions of models

```{r}
linear_reg() %>% set_engine("lm") %>% translate()
```

Back to the ames data

From before:

```{r}
library(tidymodels)
data(ames)
ames <- ames %>% mutate(Sale_Price = log10(Sale_Price))

set.seed(502)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)
```

Fitting a model:

```{r}
lm_model <- 
  linear_reg() %>% 
  set_engine("lm")

lm_form_fit <- 
  lm_model %>% 
  # Recall that Sale_Price has been pre-logged
  fit(Sale_Price ~ Longitude + Latitude, data = ames_train)

lm_xy_fit <- 
  lm_model %>% 
  fit_xy(
    x = ames_train %>% select(Longitude, Latitude),
    y = ames_train %>% pull(Sale_Price)
  )

lm_form_fit
lm_xy_fit
```

You can set arguments either at the main level, the engine level, or at the mode. Each is more specific than the last to particular packages.

```{r}
rand_forest(trees = 1000, min_n = 5) %>% 
  set_engine("ranger") %>% 
  set_mode("regression") %>% 
  translate()
```

### 6.2 Use the model results

Parsnip model objects store model outputs

Get to them via `extract_fit_engine()`

```{r}
lm_form_fit %>% extract_fit_engine()
```

This is a list lm object right?

```{r}
lm_form_fit %>% extract_fit_engine() %>% summary()
```

```{r}
model_res <- 
  lm_form_fit %>% 
  extract_fit_engine() %>% 
  summary()

# The model coefficient table is accessible via the `coef` method.
param_est <- coef(model_res)
class(param_est)
param_est
```

Use the `broom::tidy()` function to get the same output in a standardized data frame output

```{r}
tidy(lm_form_fit)
```

### 6.3 Making predictions

Things to know about `predict()`

-   The results are always a tibble

-   Column names of the tibble are always predictable

-   As many rows in the tibble as there are in the input data set

```{r}
ames_test_small <- ames_test %>% slice(1:5)
predict(lm_form_fit, new_data = ames_test_small)
```

```{r}
ames_test_small %>% 
  select(Sale_Price) %>% 
  bind_cols(predict(lm_form_fit, ames_test_small)) %>% 
  # Add 95% prediction intervals to the results:
  bind_cols(predict(lm_form_fit, ames_test_small, type = "pred_int")) 
```

Sometimes the names start with a "." to avoid conflicts with other variables

Specify the type parameter to generate things like:

-   numeric

-   class

-   prob

-   conf_int

-   pred_int

```{r}
library(tidymodels)
data(ames)
ames <- mutate(ames, Sale_Price = log10(Sale_Price))

set.seed(123)
ames_split <- initial_split(ames, prop = 0.80, strata = Sale_Price)
ames_train <- training(ames_split)
ames_test  <-  testing(ames_split)

lm_model <- linear_reg() %>% set_engine("lm")
```

### 6.4 Parsnip-adjacent packages

A list of all the models that can be used with parsnip can be found at: <https://www.tidymodels.org/find/>
